---
title: "Step3_run_wna"
author: "Iaian M"
date: '2022-07-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Params:
```{r params,eval=TRUE, echo=TRUE, message=FALSE}
root <- 'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS'
publicTemp <- 'C:/Users/Public/pspl/temp'
pgDB <- 'prov_data'

#In tables
geomTable <- 'whse.all_bc_gr_skey'
resTable <- 'whse.all_bc_res'

#Field Names
id <- 'ogc_fid'
becLbl <- 'bgc_label'
zone <- 'zone'
ele <- 'val_ele'
slp <- 'val_slp'
asp <- 'val_asp'
lat <- 'LAT_WGS84'
long <- 'LONG_WGS84'


#Out tables
latlongtbl <- 'whse.lat_long_wgs1984' 
wnaPG <- 'whse.climatewna_ogc_fid'
wnaSchema <- 'whse'
wnaTable <- 'climatewna_ogc_fid'
bpTbl <- 'pspl.biophysical_outputs'

```

## functions
```{r functions,eval=TRUE, echo=TRUE, message=FALSE}
library(sf)
library(RPostgres)
library(keyring)
library(terra)
library(rgdal)

getTableQueryPsql<-function(sql){
  conn<-dbConnect(RPostgres::Postgres(),
                  host = key_get('dbhost', keyring = 'localpsql'),
                  user = key_get('dbuser', keyring = 'localpsql'),
                  dbname = 'prov_data',
                  password = key_get('dbpass', keyring = 'localpsql'),
                  port = "5432"
  )
  on.exit(dbDisconnect(conn))
  dbGetQuery(conn, sql)}

append2Psql<-function(pgTbl,inDF){
  conn<-dbConnect(RPostgres::Postgres(),
                  host = key_get('dbhost', keyring = 'localpsql'),
                  user = key_get('dbuser', keyring = 'localpsql'),
                  dbname = 'prov_data',
                  password = key_get('dbpass', keyring = 'localpsql'),
                  port = "5432"
  )
  on.exit(dbDisconnect(conn))
  dbWriteTable(conn, pgTbl,value = inDF, overwrite = F, append = T, row.names = FALSE)}

runSQL <- function(sqlVar,sqlFile ){
          cmd <-  c("-d",pgDB,"-f",sqlFile)
          for(i in sqlVar){
            cmd <- append(cmd,"-v")
            cmd <- append(cmd,i)
          }
        print(cmd)
       print(system2("psql",args=cmd,wait=TRUE,stdout=TRUE,stderr=TRUE)) 
        }



```

## Calc lat/longs

Creating a tbale with lat/long_table coordinates to PostGres

```{r calc_lat_long,eval=FALSE, echo=FALSE, message=FALSE}
      if (grepl("\\.", latlongtbl)) {
        noSchema <- unlist(strsplit(latlongtbl, split = "[.]"))[-1]}else {
        noSchema <- latlongtbl
      }
        indexName <- paste0('inx_', noSchema)
        x1 <- paste0("latlongtbl=",latlongtbl)
        x2 <- paste0("geomTable=",geomTable)
        x3 <- paste0("indexName=",indexName)
        cmd <- c("-d",pgDB,"-f","D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/make_lat_long_fields.sql", "-v",x1, "-v", x2, "-v", x3)
        ret_code <- print(system2("psql",args=cmd,wait=TRUE,stdout=TRUE,stderr=TRUE))
        
        print('Lat long table ')
        print(getTableQueryPsql(paste0('select
                                        count(*) as rowCount,
                                        min(',long,') as xmin,
                                        min(',lat,') as ymin,
                                        max(',long,') as xmax,
                                        max(',lat,') as ymax 
                                        from ',latlongtbl,';')))
```

## Create csv WNA input tables

Exports WNA inputs table into a specified number of csv files

```{r chunk_inputs,eval=FALSE, echo=FALSE, message=FALSE}
        chunks <- 8
        rowsPerChunk <- (getTableQueryPsql(paste0('select count(*) as count from ',latlongtbl))$count/chunks) + 1
        print(rowsPerChunk)
        chunckVec <- sequence(chunks) * rowsPerChunk 
        print(chunckVec)

        exportWNAfun<-function(x){
        low <- x - rowsPerChunk
        high <- x
        sqlVar <- c(paste0("id=",id),
                    paste0("zone=",zone),
                    paste0("ele=",ele),
                    paste0("slp=",slp),
                    paste0("asp=",asp),
                    paste0("lat=",lat),
                    paste0("long=",long),
                    paste0("resTable=",resTable),
                    paste0("low=",low),
                    paste0("high=",high),
                    paste0("filename=","'", publicTemp,"/wnaInput_",x,".csv", "'"),
                    paste0("becLbl=",becLbl),
                    paste0("latlongtbl=",latlongtbl))
        runSQL(sqlVar,'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/exportWNAinputcsv.sql')
        }
        time <- system.time(lapply(chunckVec,exportWNAfun))
        print(time/60)
        
```

## Run WNA in manually

cannot run version 4.2 using command line, as a result must run manually.
good place for parralel processing when updated to new version.

## Import WNA ouputs 2 pgCSV

Imports wna outputs to pg

```{r wnaOutputs2pg,eval=FALSE, echo=FALSE, message=FALSE}
 #create a empty pg table for climatewna outputs
  sqlVar <- c(paste0("outTbl=",wnaPG))
  runSQL(sqlVar,'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/create_climatewna_pg_tbl.sql')
  
#loop through Climatewna ouputs and append to pg table
outputVec <- c('C:/Users/Public/pspl/temp/wnaInput_11849882.5_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_94799060_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_82949177.5_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_71099295_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_59249412.5_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_47399530_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_35549647.5_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_23699765_Normal_1961_1990Y.csv')
outputVec <- c('C:/Users/Public/pspl/temp/wnaInput_11849882.5_Normal_1961_1990Y.csv',
               'C:/Users/Public/pspl/temp/wnaInput_94799060_Normal_1961_1990Y.csv')

oldNames <- c('DD.0','DD.5','DD.18','DD.18.1')
newNames <- c('ddlt0','ddgt5','ddlt18','ddgt18')

appendPGwna <- function(x){
wnaR <- read.csv(file = x)
for (i in 1:length(oldNames)) {
  names(wnaR)[names(wnaR) == oldNames[i]] <- newNames[i]} 
names(wnaR) <- tolower(names(wnaR))
append2Psql(DBI::Id(schema = wnaSchema, table = wnaTable), wnaR)
}

lapply(outputVec,appendPGwna)

```
## Biophysical

Biophysical
- note check out BEC zones where biopohysical does not run...confirm this is not an error


```{r biophysicalModel,eval=TRUE, echo=FALSE, message=FALSE}


        sqlVar <- c(paste0("wnaTable=",wnaPG),
                    paste0("bptbl=",bpTbl))
        runSQL(sqlVar,'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/biophysical_no_pem.sql')

```

