---
title: "Step3_run_wna"
author: "Iaian M"
date: '2022-07-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Params:
```{r params,eval=TRUE, echo=TRUE, message=FALSE}

fileTag <- gsub(", ","",toString((floor(runif(6, min=0, max=9)))))

root <- 'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS'
inWNApath <- 'C:/Users/Public/pspl/temp'
outWNApath
pgDB <- 'prov_data'

#In tables
geomTable <- 'whse.all_bc_gr_skey'
resTable <- 'whse.all_bc_gr_skey'

#SQL files
makeLatLongSql <- "D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/make_lat_long_fields.sql"
exportInputTbls <- 'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/exportWNAinputcsv.sql'

#Field Names
id <- 'gr_skey'
becLbl <- 'bgc_label'
zone <- 'zone'
ele <- 'ele_ele'
slp <- 'slp_slp'
asp <- 'asp_asp'
lat <- 'LAT_WGS84'
long <- 'LONG_WGS84'


#Out tables
latlongtbl <- 'whse.lat_long_wgs1984' 
wnaPG <- 'whse.climatewna_ogc_fid'
wnaSchema <- 'whse'
wnaTable <- 'climatewna_ogc_fid'
bpTbl <- 'pspl.biophysical_outputs'

pgConn <- get_pg_conn_list()

```

## functions
```{r functions,eval=TRUE, echo=TRUE, message=FALSE}
library(sf)
library(faibDataManagement)
library(RPostgres)
library(keyring)
library(terra)
library(rgdal)
```

## Calc lat/longs

Creating a table with lat/long_table coordinates to PostGres

```{r calc_lat_long,eval=FALSE, echo=FALSE, message=FALSE}

##Remove schema from latlongtbl name
      if (grepl("\\.", latlongtbl)) {
        noSchema <- unlist(strsplit(latlongtbl, split = "[.]"))[-1]}else {
        noSchema <- latlongtbl
        }


indexName <- paste0('inx_', noSchema)
#Create vector of input variables for sql run
inVect <- c(paste0("latlongtbl=",latlongtbl),
            paste0("geomTable=",geomTable),
            paste0("indexName=",indexName))

#run sql file from psql  
runSQL(inVect, makeLatLongSql,pgDB)      
print('Lat long table ')
print(getTableQueryPG(paste0('select
                                        count(*) as rowCount,
                                        min(',long,') as xmin,
                                        min(',lat,') as ymin,
                                        max(',long,') as xmax,
                                        max(',lat,') as ymax 
                                        from ',latlongtbl,';'),pgConn))
```

## Create csv WNA input tables

Exports WNA inputs table into a specified number of csv files

```{r chunk_inputs,eval=FALSE, echo=FALSE, message=FALSE}
  #calculate the amount of rows to be used in each WNA input table.  (all 95 million is too large for the WNA model)
        chunks <- 8
        rowsPerChunk <- (getTableQueryPG(paste0('select count(*) as count from ',latlongtbl), pgConn)$count/chunks) + 1
        print(rowsPerChunk)
        chunckVec <- sequence(chunks) * rowsPerChunk 
        print(chunckVec)

        exportWNAfun<-function(x){
        low <- x - rowsPerChunk
        high <- x
        sqlVar <- c(paste0("id=",id),
                    paste0("zone=",zone),
                    paste0("ele=",ele),
                    paste0("slp=",slp),
                    paste0("asp=",asp),
                    paste0("lat=",lat),
                    paste0("long=",long),
                    paste0("resTable=",resTable),
                    paste0("low=",low),
                    paste0("high=",high),
                    paste0("filename=","'", inWNApath,"/wnaInput_",which(chunkVec == x),"_",fileTag,".csv", "'"),
                    paste0("becLbl=",becLbl),
                    paste0("latlongtbl=",latlongtbl))
        runSQL(sqlVar,exportInputTbls,pgDB)
        }
        time <- system.time(lapply(chunckVec,exportWNAfun))
        print(time/60)
        
```

## Run WNA in manually

cannot run version 4.2 using command line, as a result must run manually.
good place for parralel processing when updated to new version.

## Import WNA ouputs 2 pgCSV

Imports wna outputs to pg

```{r wnaOutputs2pg,eval=FALSE, echo=FALSE, message=FALSE}
 #create a empty pg table for climatewna outputs
  sqlVar <- c(paste0("outTbl=",wnaPG))
  runSQL(sqlVar,'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/create_climatewna_pg_tbl.sql',pgDB)
  
#loop through Climatewna ouputs and append to pg table
files <- list.files(path = outWNAPath, pattern = glue('99','.+Normal_1961_1990Y.csv') )

oldNames <- c('DD.0','DD.5','DD.18','DD.18.1')
newNames <- c('ddlt0','ddgt5','ddlt18','ddgt18')

appendPGwna <- function(x){
wnaR <- read.csv(file = x)
for (i in 1:length(oldNames)) {
  names(wnaR)[names(wnaR) == oldNames[i]] <- newNames[i]} 
names(wnaR) <- tolower(names(wnaR))
faibDataManagement::df2PG(DBI::Id(schema = wnaSchema, table = wnaTable), wnaR,connList,overwrite=F,append=T)
}

lapply(files,appendPGwna)

```
## Biophysical

Biophysical
- note check out BEC zones where biopohysical does not run...confirm this is not an error


```{r biophysicalModel,eval=TRUE, echo=FALSE, message=FALSE}


        sqlVar <- c(paste0("wnaTable=",wnaPG),
                    paste0("bptbl=",bpTbl))
        faibDataManagement::runSQL(sqlVar,'D:/Projects/PSPL_2022/IAIAN_VERSION/R_SCRIPTS/sql/biophysical_no_pem.sql',pgDB)

```

